---
title: "Analysis of FlowData (CSV format) "
author: "Tsuru"
date: "2022/8/16"
output: html_document
---

This script should be executed after conversion of flow data from FSC format to CSV format.
This script codes for manipulation of flow data (csv) to extract:
Mean expression levels for both wild and mutatnt types
Mutational effects
Number of cells filtered


# 0. General settings
## 0.1. Library loading
```{r General,warning=FALSE, message=FALSE}
library(ggplot2)
library(scales)
library(ggpubr)
library(gridExtra)
library(dplyr)
library(matrixStats)
library(RColorBrewer)
library(openxlsx)
library(readxl)
library(ggsci)
library(tidyverse)
library(magrittr)
library(rstatix)
library(stringr)
Sys.time()
```

## 0.2. versatile user functions
The [qw](https://ja.coder.work/so/r/240447) function in Perl.  
The [createEmptyDf](https://htsuda.net/archives/2560) function.
The [g_legend](https://github.com/hadley/ggplot2/wiki/Share-a-legend-between-two-ggplot2-graphs) function to extract legend.
```{r General}
qw <- function(...) {
  sapply(match.call()[-1], deparse)
}

createEmptyDf = function( nrow, ncol, colnames = c() ){
  data.frame( matrix( vector(), nrow, ncol, dimnames = list( c(), colnames ) ) )
}
Sys.time()
```

## 0.3. visualization
```{r General}
my_fontsize<-16
Sys.time()
```

# 1. Analysis of flow data (csv)
## 1.1. load mg1655 data and define autofluorescence and FSC gate
```{r Flow_Cytometry,warning=FALSE, message=FALSE,fig.height=3.5,fig.width=3.5,dpi=300}
dmeta<-paste(getwd(),"Meta","metadata.xlsx",sep="/") %>%
  read_xlsx()
my_volume<-"/Volumes/DailyUse"
my_path<-paste(my_volume,"Hatanaka","FlowData","CSV","MG1655.csv",sep="/")
autofl<-my_path %>% read_csv()
autofl$log10FSCpls3<-log10(autofl$FSC+1000)
# fscmode<-as.numeric(names(which.max(table(autofl$log10FSCpls3))))
hist_info <- hist(autofl$log10FSCpls3,plot=F,breaks = seq(2.5,5.5,0.01))               # Extract histogram information

# Calc. mode of FSC
fscmode<-hist_info$breaks[which(hist_info$density==max(hist_info$densit))]
paste("Mode of FSC:",fscmode) %>%
print()

# upper and lower limit of FSC gate
fsc_low<-3.675
fsc_high<-3.864

round(nrow(autofl[(autofl$log10FSCpls3>fsc_low)&(autofl$log10FSCpls3<fsc_high),])/nrow(autofl)*100) %>%
  paste(.,"% of cells remains",sep="") %>%
  print()


p1<-gghistogram(autofl,x="log10FSCpls3",y="density",fill="gray10",size=0,bins=50)+
  xlab(expression(log[10]~"FSC [a.u.]"))+
  ylab("Relative frequency [a.u.]")+
  scale_x_continuous(breaks = get_breaks(by = 1, from = 3),limits =  c(2.7, 5.5))+
  scale_y_continuous(breaks = get_breaks(by = 1, from = 0),limits =  c(0, 2.5))+
  geom_vline(xintercept = fscmode,size=0.2,linetype=2)+
  geom_vline(xintercept = c(fsc_low,fsc_high),size=0.2,linetype=2,color=get_palette("lancet",2)[2])
p1<-ggpar(p1,font.y=my_fontsize,font.x=my_fontsize,font.xtickslab = my_fontsize,font.ytickslab = my_fontsize)
p1

# Define autofluorescence
autofl.th<-quantile(autofl[(autofl$log10FSCpls3>fsc_low)&(autofl$log10FSCpls3<fsc_high),]$FITC,0.995) 

p2<-gghistogram(autofl,x="FITC",y="density",fill = "gray10", size=0,bins=50,
            xlab="Green FI [a.u.]",
            ylab="Relative frequency [a.u.]")+
  scale_x_continuous(breaks = get_breaks(by = 500, from = 0),limits =  c(-300, 1100))+
  scale_y_continuous(breaks = get_breaks(by = 0.002, from = 0),limits =  c(0, 0.0045))+
  geom_vline(xintercept = autofl.th,size=0.2,linetype=2,color=get_palette("lancet",2)[2])
p2<-ggpar(p2,font.y=my_fontsize,font.x=my_fontsize,font.xtickslab = my_fontsize,font.ytickslab = my_fontsize)
p2

rm(autofl)
Sys.time()
```

## 1.2. density plot of Green FI for all strains
```{r Flow_Cytometry,warning=FALSE, message=FALSE,fig.height=2,fig.width=16,dpi=300}
my_tmp<-dmeta[(!is.na(dmeta$Used))&(!is.na(dmeta$Genotype))&(!is.na(dmeta$Type)),]
mylist<-unique(my_tmp[(my_tmp$Used=="Yes")&(my_tmp$Genotype=="wild type")&(my_tmp$Type=="Promoter"),]$Promoter_Name)
my_df<-data.frame(Promoter_Name=mylist,WT_exprs=NA,MT_exprs=NA,
                  WT_sd=NA,MT_sd=NA,
                  logFC=NA,WT_NbCells=NA,MT_NbCells=NA)
figlist<-list()
for(i in 1:length(mylist)){
  
  ddw<-dmeta[(dmeta$Promoter_Name %in% mylist[i])&(dmeta$Genotype=="wild type"),]$CSV_filename_new
  ddm<-dmeta[(dmeta$Promoter_Name %in% mylist[i])&(dmeta$Genotype=="mutant library"),]$CSV_filename_new
  dw<-paste(my_volume,"Hatanaka","FlowData","CSV",ddw,sep="/") %>%
    read_csv()
  dw$Name<-"WT"
  dm<-paste(my_volume,"Hatanaka","FlowData","CSV",ddm,sep="/") %>%
    read_csv()
  dm$Name<-"Mut"
  dwm<-bind_rows(dm,dw)
  dwm<-dwm[dwm$FSC>-999,]
  dwm$log10FSCpls3<-log10(dwm$FSC+1000)
  dwm$FITCsbt<-dwm$FITC-autofl.th
  dwm$log10FITCnorm<-ifelse(dwm$FITCsbt>0,log10(dwm$FITCsbt),NA)
  dwm<-dwm[!is.na(dwm$log10FITCnorm),]
  dwm<-dwm[(dwm$log10FSCpls3>fsc_low)&(dwm$log10FSCpls3<fsc_high)&(dwm$PE<1000),]
  dwm$Name<-factor(dwm$Name,levels=qw(WT,Mut))
  
  my_df$WT_exprs[i]<-mean(dwm[dwm$Name=="WT",]$log10FITCnorm)
  my_df$MT_exprs[i]<-mean(dwm[dwm$Name=="Mut",]$log10FITCnorm)
  my_df$WT_sd[i]<-sd(dwm[dwm$Name=="WT",]$log10FITCnorm)
  my_df$MT_sd[i]<-sd(dwm[dwm$Name=="Mut",]$log10FITCnorm)
  my_df$logFC[i]<-my_df$MT_exprs[i]-my_df$WT_exprs[i]
  my_df$WT_NbCells[i]<-nrow(dwm[dwm$Name=="WT",])
  my_df$MT_NbCells[i]<-nrow(dwm[dwm$Name=="Mut",])
  
  g<-gghistogram(dwm,x="log10FITCnorm",y="density", size=0,bins=50,alpha=0.4,
            legend="",legend.title="",
            # add="mean", add.params = list(size=0.5,linetype=5),
            color="Name",fill="Name",palette = c("#00AFBB","#D870AD"),
            xlab="",ylab="",subtitle=mylist[i]
            # xlab=expression(log[10]~"Green FI [a.u.]"),
            # ylab="Relative frequency [a.u.]"
            )+
    # annotate("text",x=-Inf,y=Inf,label= mylist[i],hjust = -0.2,vjust = 1,size=5)+
    scale_x_continuous(breaks = get_breaks(by = 1, from = 1),limits =  c(1, 5))
    
    my_gb<-ggplot_build(g)
    my_max<-max(my_gb$data[[1]]$density)
    if(my_max<2){
      g<-g+scale_y_continuous(breaks = get_breaks(by = 1, from = 0),limits = c(0,2))
    }else{
      g<-g+scale_y_continuous(breaks = get_breaks(by = 1, from = 0))
    }
  g<-ggpar(g,font.tickslab =c(15),font.subtitle = c(20))
  figlist[[i]]<-g
}

rm(dw,dm,dwm,g)

# k<-1
# for(l in 1:8){
#   if(l<=7){
#     grid.arrange(figlist[[k+1]],figlist[[k+2]],figlist[[k+3]],figlist[[k+4]],figlist[[k+5]],figlist[[k+6]],figlist[[k+7]],figlist[[k+8]],figlist[[k+9]],figlist[[k+10]],nrow=1)
#     k<-k+10
#   }else{
#     grid.arrange(figlist[[1]],figlist[[2]],ncol=10)
#   }
# }

k<-0
for(l in 1:9){
    grid.arrange(figlist[[k+1]],figlist[[k+2]],figlist[[k+3]],figlist[[k+4]],figlist[[k+5]],figlist[[k+6]],figlist[[k+7]],figlist[[k+8]],nrow=1)
    k<-k+8
}
rm(figlist)
Sys.time()
```
## Output
```{r Generic_prop,warning=FALSE, message=FALSE,fig.height=3,fig.width=3,dpi=300}
# write_csv(x = my_df,file = paste(getwd(),"Stats_of_FlowData","flowstats.csv",sep="/"))
Sys.time()
```


## Check
```{r Generic_prop,warning=FALSE, message=FALSE,fig.height=3,fig.width=3,dpi=300}
ggscatter(my_df,x="WT_exprs",y="MT_exprs",alpha=0.5)+
  geom_abline(slope = 1,intercept = 0)
Sys.time()
```
