---
title: "RMFmodel"
output: html_document
date: "2022-08-04"
---
# 0. General Settings
## 0.1. Workspace
### Load workspace
```{r workspace_load,warning=FALSE, message=FALSE}
# load("./RMF.RData")
```
### Save workspace
```{r workspace_save,warning=FALSE, message=FALSE}
# save.image("./RMF.RData")
```
## 0.2. Library loading
```{r library,warning=FALSE, message=FALSE}
library(viridis)
library(ggplot2)
library(GGally)
library(scales)
library(ggpubr)
library(grid)
library(gtable)
library(gridExtra)
library(dplyr)
library(matrixStats)
library(preprocessCore)
library(RColorBrewer)
library(openxlsx)
library(ggsci)
library(cowplot)
library(corrplot)
library(Hmisc)
# https://stackoverflow.com/questions/59419647/hmisc-package-or-namespace-failed-to-load-no-package-called-latticeextra
library(corrplot)
library(tidyverse)
library(magrittr)
library(dendextend)
library(FactoMineR)
library(factoextra)
library(proxy)
library(rstatix)
# library(plot3D)
# library(clusterProfiler)
# library(pathview)
# library(org.EcK12.eg.db)
# library(rvest)
# library(corrr)
# library(GOSemSim)
# library(rrvgo)
# library(qdapTools)
# library(AnnotationDbi)
# library(ggVennDiagram)
# library(rJava)
# library(venneuler)
library(gtools)
library(parallel)
```
## 0.3. versatile user functions
The [qw](https://ja.coder.work/so/r/240447) function in Perl.  
The [createEmptyDf](https://htsuda.net/archives/2560) function.
The [g_legend](https://github.com/hadley/ggplot2/wiki/Share-a-legend-between-two-ggplot2-graphs) function to extract legend.
```{r functions}
qw <- function(...) {
  sapply(match.call()[-1], deparse)
}

createEmptyDf = function( nrow, ncol, colnames = c() ){
  data.frame( matrix( vector(), nrow, ncol, dimnames = list( c(), colnames ) ) )
}

g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

pValue_sp_text=function(dtf){
  pcor_sp<-cor.test(x=dtf[,1],y=dtf[,2],method="spearman")
  rho_sp<-round(pcor_sp$estimate,digits = 2)
  pv_sp<-ifelse(pcor_sp$p.value<0.05," P<0.05",paste(" P==",round(pcor_sp$p.value,digits = 1),sep = ""))
  rhopvtxt_sp<-paste("list(italic(rho)==", rho_sp,", ",pv_sp,")")
  return(rhopvtxt_sp)
}

pValue_pr_text=function(dtf){
  pcor_pr<-cor.test(x=dtf[,1],y=dtf[,2],method="pearson")
  r_pr<-round(pcor_pr$estimate,digits = 2)
  pv_pr<-ifelse(pcor_pr$p.value<0.05," P<0.05",paste(" P==",round(pcor_pr$p.value,digits = 1),sep = ""))
  rpvtxt_pr<-paste("list(italic(r)==", r_pr,", ",pv_pr,")")
  return(rpvtxt_pr)
}
```

## create a binary matrix
```{r Data_Source,warning=FALSE, message=FALSE,fig.height=5,fig.width=5,dpi=300}
N<-10 # number of genetic sites
c<-1
gt_lst<-permutations(2,N,v=c(0,1),repeats.allowed=TRUE)
head(gt_lst)
length(unique(do.call(paste,c(as.data.frame(gt_lst),sep="_"))))
```
## 0.4. user functions
### 0.4.1. calculation of hamming distance
```{r Data_Source,warning=FALSE, message=FALSE,fig.height=5,fig.width=5,dpi=300}
hmd<-function(x,y){
  x1<-matrix(rep(x,each=nrow(y)),ncol=ncol(y))
  y1<-abs(y-x1)
  y_out<-rowSums(y1)
  return(y_out)
}
temp<-floor(runif(N,0,2))
hmd(temp,gt_lst)
```
### 0.4.2. calculate mutation effect
calculate mean expression level of mutants whose genotypes are defferent by single mutations from wild type
```{r Data_Source,warning=FALSE, message=FALSE,fig.height=1,fig.width=1,dpi=300}
mut_effect<-function(gt_lst,expression){
  xout<-list()
  z<-list()
  for(i in 1:nrow(gt_lst)){
    x<-hmd(gt_lst[i,],gt_lst)
    y1<-expression[which(x==1)]
    y2<-expression[which(x==1)]-expression[i]
    xout[[i]]<-mean(y2,na.rm=T)
    z[[i]]<-paste(sum(sign(y2)>0),sum(sign(y2)==0),sum(sign(y2)<0),sep=",")
    # z[[i]]<-ifelse(sum(sign(y2))==length(y2),"sink",ifelse(sum(sign(y2))==-length(y2),"peak","composite"))
  }
  xout<-unlist(xout)
  zout<-unlist(z)
  return(list(xout,zout))
}

uphill_ratio<-function(array){
  x_out<-list()
  for(ii in 1:length(array)){
    x<-as.numeric(str_split(array[ii],",")[[1]])
    x_out[[ii]]<-x[1]/sum(x)
  }
  return(unlist(x_out))
}

# temp<-RMF_exlandscape_list[[2]]
# ggscatter(temp,x="hmdist",y="expression",size=0.3,alpha=0.1)
# i<-1
# 
# temp$expression[which(hmd(gt_lst[i,],gt_lst)==1)]
# temp$expression[which(hmd(gt_lst[i,],gt_lst)==0)]
# temp1<-temp$expression[which(hmd(gt_lst[i,],gt_lst)==1)]-temp$expression[which(hmd(gt_lst[i,],gt_lst)==0)]
# mut_effect(gt_lst,temp$expression)
```

# 1. Expression landscape 
create expression landscape
## 1.1. RMF
https://github.com/song88180/fitness-landscape-error/blob/master/1_Generate_Fitness_Landscape/RMF_model_generation.ipynb
https://colorate.azurewebsites.net/ja
```{r Data_Source,warning=FALSE, message=FALSE,fig.height=5,fig.width=5,dpi=300}
Nland<-10
std<-1 # 0.3 for modified RMF
RMF_exlandscape<-function(gt_lst,std,Nland){
  x<-gt_lst
  y<-Nland
  xout<-list()
  for (i in 1:y){
    wt<-floor(runif(N,0,2))
    hmdz<-hmd(wt,x)
    z <- -c*hmdz+rnorm(2^N, mean=0, sd=std) # Normal RMF
    # z <- -c*hmdz+rnorm(2^N, mean=0, sd=std*hmdz^0.6) #<-------------------------- modified RMF
    z1<-min(z)
    z2<-max(z)
    z <-(z-z1)/(z2-z1)
    x_max<-x[which(z==max(z)),]
    hmdz_norm<-hmd(x_max,x)
    me<-mut_effect(x,z)
    z<-data.frame(genotype=do.call(paste,c(as.data.frame(x),sep="_")),expression=z,muteffect=me[[1]],shape=me[[2]],hmdist=hmdz_norm)
    z$genotype<-as.character(z$genotype)
    xout[[i]]<-z
    
  }
  return(xout)
}

c=1
RMF_exlandscape_list<-RMF_exlandscape(gt_lst,std,Nland)

temp1<-RMF_exlandscape_list[[4]]
temp1$uphillratio<-uphill_ratio(temp1$shape)
ggscatter(temp1,x="hmdist",y="expression",color="uphillratio",alpha=0.5,size=2,
          xlab="Hamming distance from the fittest",ylab="Expression level",legend.title="")+
  scale_color_gradient2(low  = "#C26B51",mid  = "#DDCE47",high = "#529985",midpoint = 0.5,limits=c(0,1))

ggscatter(temp1,x="expression",y="muteffect",color="uphillratio",alpha=0.5,size=2,
          xlab="Expression level",ylab="Mutational effects",legend.title="")+
  scale_color_gradient2(low  = "#C26B51",mid  = "#DDCE47",high = "#529985",midpoint = 0.5)
```
## 1.2. modified NK model
```{r Data_Source,warning=FALSE, message=FALSE,fig.height=5,fig.width=5,dpi=300}
imatrix_rand<-function(N,K){
  Int_matrix_rand<-matrix(0,N,N)
  for(aa1 in 1:N){
    Indexes_1<-seq(1,N)
    Indexes_1[aa1]<-NA
    Indexes_1<-Indexes_1[!is.na(Indexes_1)]
    Indexes_1<-c(sample(Indexes_1),aa1)
    Chosen_ones<-Indexes_1[c(N-(K)):N]   # this takes the last K+1 indexes
    for(aa2 in Chosen_ones){
      Int_matrix_rand[aa1,aa2]<-1 # we turn on the interactions with K other variables
    }
  }
  return(Int_matrix_rand)
}

calc_fit<-function(NK_land_, inter_m, Current_position, Power_key_){
  # Fit_vector<-matrix(0,nrow = N,ncol=1)
  Fit_vector<-rep(0,N)
  for(ad1 in 1:N){
    Fit_vector[ad1]<-NK_land_[1+sum(Current_position*inter_m[ad1,]*Power_key_),ad1]
  }
  return(Fit_vector)
}

comb_and_values<-function(gt_lst,NK_land_,Power_key_,inter_m){
  Comb_and_value<-matrix(0,nrow=2^N,ncol = N+1)
  c1<-1
  pmt<-gt_lst
  for(c2 in 1:nrow(pmt)){
    Combination1<-pmt[c2,]
    fit_1<-calc_fit(NK_land_,inter_m,Combination1,Power_key_)
    Comb_and_value[c1,1:N]<-Combination1
    Comb_and_value[c1,N+1]<-mean(fit_1)
    c1<-c1+1
  }
  return(Comb_and_value)
}

normalize_nk<-function(array){
  max_nk<-max(array)
  min_nk<-min(array)
  return((array-min_nk)/(max_nk-min_nk))
}

modify_nk<-function(array){
  # array<-abs(array-0.5)
  array<-array-0.5
  array[array<0]<-0
  return(normalize_nk(array))
}

NK_exlandscape<-function(gt_lst,Nland,modify=TRUE){
  Power_key<-2^rev(seq(0,N-1))
  xout<-list()
  for(ii in 1:Nland){
    Int_matrix<-imatrix_rand(N,K)
    NK_land<-matrix(runif((2^N)*N),ncol=N)
    x<-comb_and_values(gt_lst,NK_land, Power_key, Int_matrix)
    x<-data.frame(genotype=do.call(paste,c(as.data.frame(x[,1:N]),sep="_")),expression=x[,N+1])
    x$genotype<-as.character(x$genotype)
    x$expression<-normalize_nk(x$expression)
    if(modify){
      x$expression<-modify_nk(x$expression)
    }
    gt_max<-gt_lst[which(x$expression==max(x$expression)),]
    x$hmdist<-hmd(gt_max,gt_lst)
    y<-mut_effect(gt_lst,x$expression)
    x$muteffect<-y[[1]]
    x$shape<-y[[2]]
    xout[[ii]]<-x
  }
  return(xout)
}
# library(paletteer)
# library(ggthemes)

# https://r-charts.com/color-palettes/

N<-10
K<-3
Nland<-10
NK_exlandscape_lst<-NK_exlandscape(gt_lst,Nland,modify=FALSE)
temp1<-NK_exlandscape_lst[[1]]
temp1$uphillratio<-uphill_ratio(temp1$shape)
ggscatter(temp1,x="expression",y="muteffect",color="uphillratio",alpha=0.5,size=2,
          xlab="expression level",ylab="mutation effect",legend.title="")+
  scale_color_gradient2(low  = "#C26B51",mid  = "#DDCE47",high = "#529985",midpoint = 0.5,limits=c(0,1))

ggscatter(temp1,x="hmdist",y="expression",color="uphillratio",alpha=0.5,size=2,
          xlab="Hamming distance from the fittest",ylab="mutation effect",legend.title="")+
  scale_color_gradient2(low  = "#C26B51",mid  = "#DDCE47",high = "#529985",midpoint = 0.5,limits=c(0,1))

# ggscatter(temp,x="hmdist",y="expression",alpha=0.5,color="shape",size=0.5)
```
## 1.3. generate initial genotypes whose expression levels are at the middle levels 
```{r Data_Source,warning=FALSE, message=FALSE,fig.height=5,fig.width=5,dpi=300}
exlevel<-0.3
initial_genotypes<-function(gt_lst,exlandscape_list,exlevel){
  xout<-list()
  for(ii in 1:length(exlandscape_list)){
    x<-exlandscape_list[[ii]]
    xout[[ii]]<-x[x$expression>exlevel,]$genotype
  }
  return(xout)
}
temp<-initial_genotypes(gt_lst,RMF_exlandscape_list,exlevel)
sample(x=temp[[1]],size=10)
```

# 2. Fitness landscape
create fitness landscape
```{r Data_Source,warning=FALSE, message=FALSE,fig.height=5,fig.width=5,dpi=300}
fitlandscape<-function(RMF_exlandscape_list,intrcpt){
  k<-1
  x<-list()
  y<-intrcpt
  z<-RMF_exlandscape_list
  for(i in 1:length(y)){
    for(j in 1:length(z)){
      xi<-z[[j]]
      xi$intercept<-y[i]
      xi$fitness<-(1-y[i])*xi$expression+y[i]
      x[[k]]<-xi
      k<-k+1
    }
  }
  return(x)
}
intrcpt<-c(0.3,0.5,0.99)
RMF_landscape_list<-fitlandscape(RMF_exlandscape_list,intrcpt)
head(RMF_landscape_list[[1]])
```
# 3. Mutation matrix 
## 3.1. parameters
```{r Data_Source,warning=FALSE, message=FALSE,fig.height=5,fig.width=5,dpi=300}
# n = len(next(iter(fitness)))
mu <- 0.001#1e-4 # mutation rate as in Szendro et al. 2013 
m <- nrow(gt_lst) # Number of genotypes
```
## 3.2. create Mutation matrix
```{r Data_Source,warning=FALSE, message=FALSE,fig.height=1,fig.width=1,dpi=300}
mutation_matrix<-function(genotypes, mu){
    m <- nrow(genotypes)
    M <- diag(1.0, m)
    for(i in 1:m){
      g1<-hmd(genotypes[i,],genotypes)
      M[which(g1==1),i]<-mu
      M[i,i]<-M[i,i]-length(which(g1==1))*mu
    }
    return(M)
}
M<-mutation_matrix(gt_lst,mu)
l<-ifelse(nrow(M)>32,32,nrow(M))
corrplot(M[1:l,1:l],cl.lim=c(0,1), is.corr = FALSE,tl.cex = 0.3,cl.cex = 0.3,mar=c(1,0,1,0),tl.col = "black", method = "color")
```
# 4. Selection matrix
## 4.1. create Selection matrix
```{r Data_Source,warning=FALSE, message=FALSE,fig.height=1,fig.width=1,dpi=300}
selection_matrix<-function(fitness){
    return(diag(fitness, length(fitness)))
}
S<-selection_matrix(RMF_landscape_list[[1]]$fitness)
l<-ifelse(nrow(S)>32,32,nrow(S))
corrplot(S[1:l,1:l],cl.lim=c(0,1), is.corr = FALSE,tl.cex = 0.3,cl.cex = 0.3,mar=c(1,0,1,0),tl.col = "black", method = "color")
```
## 4.2. create single matrix for mutation and the linear part of selection
```{r Data_Source,warning=FALSE, message=FALSE,fig.height=1,fig.width=1,dpi=300}
E<-M %*% S 
l<-ifelse(nrow(E)>32,32,nrow(E))
corrplot(E[1:l,1:l],cl.lim=c(0,1), is.corr = FALSE,tl.cex = 0.3,cl.cex = 0.3,mar=c(1,0,1,0),tl.col = "black", method = "color")
```
# 5. Simulate evolution (RMF)
## 5.1. adaptation (RMF)
create a function to simulate evolution based on a given landscape
```{r RMF,warning=FALSE, message=FALSE,fig.height=2,fig.width=2,dpi=300}
Npop<-1e9
generations<-50 # 2^12
intrcpt<-c(0.1,0.4,0.7)
N<-5 # number of genetic sites
c<-1
std<-2
Nland<-10
gt_lst<-permutations(2,N,v=c(0,1),repeats.allowed=TRUE)
RMF_exlandscape_list<-RMF_exlandscape(gt_lst,std,Nland)
RMF_landscape_list<-fitlandscape(RMF_exlandscape_list,intrcpt)

landscape_list<-RMF_landscape_list[[12]]
# gstart<-landscape_list$genotype[which(landscape_list$fitness==min(landscape_list$fitness))] # activate when using maladaptive genotype
gstart<-paste(floor(runif(N,0,2)),collapse="_") # activate when using maladaptive genotype
gend<-landscape_list$genotype[which(landscape_list$fitness==max(landscape_list$fitness))]
mu <- 0.0001#1e-4 # mutation rate as in Szendro et al. 2013 
m <- nrow(gt_lst) # of genotypes
M<-mutation_matrix(gt_lst,mu)
S<-selection_matrix(landscape_list$fitness)
E<-M %*% S

# gstart<-paste(c(0,0,0,0,1),collapse = "_")
# gend<-paste(c(1,1,1,1,1),collapse = "_")
adaptation<-function(landscape_list,gstart,gend,verbose){
  g0i<-which(landscape_list$genotype==gstart)
  if(verbose){
    print(paste("starting genotype:", landscape_list$genotype[g0i]))
    print(paste("starting fitness:", landscape_list$fitness[g0i]))
  }
  g1i<-which(landscape_list$genotype==gend)
  
  #p is the frequencies vector
  p<-matrix(NA, nrow = nrow(landscape_list), ncol = generations)
  p[,1]<-0
  p[g0i,1]<-1.0
  
  # pom holds the index of the most frequent genotype at each time point
  pom<-matrix(NA, nrow = nrow(landscape_list), ncol = 1)
  pom[1]<-which(p[,1]==max(p[,1]))
  
  # W holds the population mean fitness at each time point
  W <- matrix(NA, nrow = generations, ncol = 1)
  W[1] = diag(S) %*% p[,1]  # mean fitness calculation
  
  for (k in 2:generations){
    if((verbose)&(k %% c(1000) == 0)){
      print(k, end=" ")
    }
    p[,k] <- E %*% p[, k - 1]  # selection and mutation are deterministic
    p[,k] <- p[, k]/sum(p[,k])  # normalize frequencies vector
    p[,k] <- unlist(lapply(p[,k],rbinom,n=1,size =Npop))/Npop  # random genetic drift is stochastic

    W[k] <- diag(S) %*% p[,k]   # mean fitness calculation
    pom[k]<-which(p[,k]==max(p[,k]))
    
    # break if reached fixation
    if(p[g1i,k] > 0.99){ #<---------------------------should be adjusted according to mu
        break
    }
  }
  # return(list(k, p, W,pom[k]))
  return(pom[k])
}


res <- adaptation(landscape_list,gstart,gend,verbose=FALSE)
# landscape_list$expression[[res[[4]]]]
# landscape_list$muteffect[[res[[4]]]]
# landscape_list$shape[[res[[4]]]]

```

## 5.2. Simulation of evolution based on the RMF model
Simulate evolution for various landscapes and calc. mutational effect
```{r RMF,warning=FALSE, message=FALSE,fig.height=2,fig.width=2,dpi=300}
Npop<-1e5
generations<-30 # 2^12
intrcpt<-c(0.3,0.5,0.99)
N<-10 # number of genetic sites
c<-1 # 1 for RMF, 0 for HoC 
std<-1 # 0.3 for modified RMF, 1 for RMF
Nland<-10
mu <- 0.001#1e-4 # mutation rate as in Szendro et al. 2013 
exlevel<--0.1
# Nini<-100 # activate if start from random genotypes

gt_lst<-permutations(2,N,v=c(0,1),repeats.allowed=TRUE)
m <- nrow(gt_lst) # of genotypes
RMF_exlandscape_list<-RMF_exlandscape(gt_lst,std,Nland)
RMF_landscape_list<-fitlandscape(RMF_exlandscape_list,intrcpt)
M<-mutation_matrix(gt_lst,mu)

gt_initials<-initial_genotypes(gt_lst,RMF_exlandscape_list,exlevel)
gstart_lst<-list()
for(i in 1:length(gt_initials)){
  temp<-gt_initials[[i]]
  Nini<-min(c(length(temp),600))
  gstart_lst[[i]]<-sample(x=temp,size=Nini)
}

for (i in 1:length(RMF_landscape_list)){
  landscape_list<-RMF_landscape_list[[i]]
  gend<-landscape_list$genotype[which(landscape_list$fitness==max(landscape_list$fitness))]
  S<-selection_matrix(landscape_list$fitness)
  E<-M %*% S
  l<-ifelse(i%%Nland==0,Nland,i%%Nland)
  gstart<-gstart_lst[[l]]
  res<-mclapply(gstart,adaptation,landscape_list = landscape_list,gend = gend,verbose=FALSE)
  res<-unlist(res)
  temp<-createEmptyDf(nrow=length(res),ncol=4,colnames=qw(intrcpt,expression,muteffect,shape))
  temp$intrcpt<-landscape_list$intercept[1]
  temp$expression<-landscape_list$expression[res]
  temp$muteffect<-landscape_list$muteffect[res]
  temp$shape<-landscape_list$shape[res]
  if(i==1){
    dMutEffect<-temp
  }else{
    dMutEffect<-rbind(dMutEffect,temp)
  }
  print(i)
}
```




## 5.3. Visualization
### 5.3.1.
```{r RMF,warning=FALSE, message=FALSE,fig.height=3.2,fig.width=9.6,dpi=300}
dMutEffect$intrcpt<-factor(dMutEffect$intrcpt,levels=intrcpt)
temp<-get_palette("lancet",2)
temp<-c(temp[c(2,1)],"gray40")
myslope<--1
myintercept<-0.5
ggscatter(dMutEffect,x="expression",y="muteffect",color = "intrcpt",alpha=0.2,size=0.2)

temp0<-dMutEffect
tempx<-list()
figlist<-list()
for(i in 1:length(intrcpt)){
  tempx[[i]]<-temp0[temp0$intrcpt==intrcpt[i],]
  tempx[[i]]<-tempx[[i]][order(tempx[[i]]$expression),]
  tempx[[i]]$Runmed<-runmed(tempx[[i]]$muteffect,101)
  g<-ggscatter(tempx[[i]],x="expression",y="muteffect",color = temp[i],alpha=0.1,size=0.2,xlim=c(0,1),ylim=c(-0.4,0.4),
               xlab="Expression level [a.u.]", ylab="Mutational effect [a.u.]")+
    scale_x_continuous(breaks = get_breaks(by = 0.5, from = 0),limits =  c(0, 1)) +
    geom_line(aes(x=expression,y=Runmed),color = temp[i])+
    annotate("text",x=Inf,y=Inf,label=paste("Fb=",intrcpt[i],sep=" "), hjust = 1,vjust = 1,size=7)
  g<-ggpar(g,font.xtickslab = c(18),font.ytickslab = c(18),font.x = c(18),font.y=c(18))
  # g<-g+geom_abline(slope = myslope,intercept = myintercept,size=0.2)+geom_vline(xintercept = c(0.5,0.75),size=0.2)
  figlist[[i]]<-g
}
grid.arrange(figlist[[1]], figlist[[2]], figlist[[3]], nrow = 1)

for(i in 1:length(intrcpt)){
  if(i==1){
    dfrunmed<-tempx[[i]]
  }else{
    dfrunmed<-rbind(dfrunmed,tempx[[i]])
  }
}

```
### 5.3.2.
```{r RMF,warning=FALSE, message=FALSE,fig.height=3.2,fig.width=9.6,dpi=300}
temp1<-RMF_exlandscape_list[[1]]
temp1$uphillratio<-uphill_ratio(temp1$shape)

g1<-ggscatter(temp1,x="hmdist",y="expression",color="uphillratio",alpha=0.5,size=2,
          xlab="Hamming distance",ylab="Expression level [a.u]",legend="")+
    scale_x_continuous(breaks = get_breaks(by = 2, from = 0),limits =  c(0, 10)) +
  scale_y_continuous(breaks = get_breaks(by = 0.5, from = 0),limits =  c(0, 1)) +
  scale_color_gradient2(low  = "#C26B51",mid  = "#DDCE47",high = "#529985",midpoint = 0.5,limits=c(0,1))
g1<-ggpar(g1,font.xtickslab = c(18),font.ytickslab = c(18),font.x = c(18),font.y=c(18))

temp3<-RMF_landscape_list[[1]]
temp3$uphillratio<-uphill_ratio(temp3$shape)
g2<-ggscatter(temp3,x="hmdist",y="fitness",color="uphillratio",alpha=0.5,size=2,ylim=c(0,1),
          xlab="Hamming distance",ylab="Fitness [a.u]",legend="")+
  geom_hline(yintercept = intrcpt[1],color=temp[1])+
  scale_x_continuous(breaks = get_breaks(by = 2, from = 0),limits =  c(0, 10)) +
  scale_y_continuous(breaks = get_breaks(by = 0.2, from = 0),limits =  c(0, 1)) +
  scale_color_gradient2(low  = "#C26B51",mid  = "#DDCE47",high = "#529985",midpoint = 0.5,limits=c(0,1))
g2<-ggpar(g2,font.xtickslab = c(18),font.ytickslab = c(18),font.x = c(18),font.y=c(18))

g3<-ggscatter(temp1,x="expression",y="muteffect",color="uphillratio",alpha=0.5,size=2,
          xlab="Expression level [a.u.]",ylab="Mutational effect [a.u.]",legend="",xlim=c(0,1),ylim=c(-0.4,0.4))+
          scale_x_continuous(breaks = get_breaks(by = 0.5, from = 0),limits =  c(0, 1)) +
          scale_color_gradient2(low  = "#C26B51",mid  = "#DDCE47",high = "#529985",midpoint = 0.5,limits=c(0,1))
g3<-ggpar(g3,font.xtickslab = c(18),font.ytickslab = c(18),font.x = c(18),font.y=c(18))

grid.arrange(g1,g3,g2, nrow = 1)

p_led<-ggscatter(temp1,x="expression",y="muteffect",color="uphillratio",alpha=0.5,size=2,
          xlab="Expression level [a.u.]",ylab="Mutational effect [a.u.]",legend="right",legend.title="",xlim=c(0,1),ylim=c(-0.4,0.4))+
          scale_x_continuous(breaks = get_breaks(by = 0.5, from = 0),limits =  c(0, 1)) +
          scale_color_gradient2(low  = "#C26B51",mid  = "#DDCE47",high = "#529985",midpoint = 0.5,limits=c(0,1),breaks=c(0,0.5,1))
q<-ggpar(p_led,font.legend = 16)
leg <- get_legend(q)
as_ggplot(leg)
```
### 5.3.3.
```{r RMF,warning=FALSE, message=FALSE,fig.height=3.2,fig.width=3.2,dpi=300}
temp2<-data.frame(expression=rep(c(0,1),length(intrcpt)),intrcpt=rep(intrcpt,each=2))
temp2$fitness<-(1-temp2$intrcpt)*temp2$expression+temp2$intrcpt
temp2$intrcpt<-factor(temp2$intrcpt,levels=intrcpt)
g4<-ggline(temp2,x="expression",y="fitness",color="intrcpt",palette=temp,
       numeric.x.axis = TRUE,plot_type="l",xlim=c(0,1),ylim=c(0,1),
           xlab="Expression level [a.u.]", ylab="Fitness [a.u.]",legend="")+
  scale_x_continuous(breaks = get_breaks(by = 0.2, from = 0),limits =  c(0, 1)) +
  scale_y_continuous(breaks = get_breaks(by = 0.2, from = 0),limits =  c(0, 1))

g5<-ggline(dfrunmed,x="expression",y="Runmed",color = "intrcpt",palette = temp,
           numeric.x.axis = TRUE,plot_type="l",xlim=c(0,1),ylim=c(-0.4,0.4),
           xlab="Expression level [a.u.]", ylab="Mutational effect [a.u.]",legend="")+
  scale_x_continuous(breaks = get_breaks(by = 0.5, from = 0),limits =  c(0, 1)) +
  scale_y_continuous(breaks = get_breaks(by = 0.2, from = -0.4),limits =  c(-0.4, 0.4))
g5<-ggpar(g5,font.xtickslab = c(18),font.ytickslab = c(18),font.x = c(18),font.y=c(18))

# g1<-g1+geom_abline(slope = myslope,intercept = myintercept,size=0.2)
g5.1<-ggline(dfrunmed,x="expression",y="Runmed",color = "intrcpt",palette = temp,
           numeric.x.axis = TRUE,plot_type="l",xlim=c(0.6,0.8),ylim=c(-0.2,0),
           xlab="Expression level [a.u.]", ylab="Mutational effect [a.u.]")
g5
```
## 5.4. Output
```{r RMF,warning=FALSE, message=FALSE,fig.height=3.2,fig.width=3.2,dpi=300}
# expression landscape
for(i in 1:length(RMF_exlandscape_list)){
  my_tmp<-RMF_exlandscape_list[[i]]
  my_tmp$ID<-i
  if(i==1){
    my_df<-my_tmp
  }else{
    my_df<-bind_rows(my_df,my_tmp)
  }
}
my_df<-my_df %>% select(ID,everything())
write_csv(my_df,file = paste(getwd(),"Output","ExprsLandscape_RMF.csv",sep = "/"))

# fitness landscape
my_ids<-rep(seq(1,length(RMF_exlandscape_list)),each=length(intrcpt))
for(i in 1:length(RMF_landscape_list)){
  my_tmp<-RMF_landscape_list[[i]]
  my_tmp$ID<-my_ids[i]
  if(i==1){
    my_df<-my_tmp
  }else{
    my_df<-bind_rows(my_df,my_tmp)
  }
}
my_df<-my_df %>% select(ID,everything())
write_csv(my_df,file = paste(getwd(),"Output","FitnessLandscape_RMF.csv",sep = "/"))

# mutational effect
write_csv(dMutEffect,file = paste(getwd(),"Output","MutEffect_RMF.csv",sep = "/"))
```
# 6. Simulate evolution (modified NK)
## 6.1. adaptation (modified NK)
```{r Data_Source,warning=FALSE, message=FALSE,fig.height=2,fig.width=2,dpi=300}
Npop<-1e9
generations<-30 # 2^12
intrcpt<-c(0.3,0.5,0.99)
N<-10 # number of genetic sites
# c<-1
# std<-2
Nland<-10
K<-5
gt_lst<-permutations(2,N,v=c(0,1),repeats.allowed=TRUE)
NK_exlandscape_list<-NK_exlandscape(gt_lst,Nland,modify=FALSE)
NK_landscape_list<-fitlandscape(NK_exlandscape_list,intrcpt)

landscape_list<-NK_landscape_list[[12]]
# gstart<-landscape_list$genotype[which(landscape_list$fitness==min(landscape_list$fitness))] # activate when using maladaptive genotype
gstart<-paste(floor(runif(N,0,2)),collapse="_") # activate when using maladaptive genotype
gend<-landscape_list$genotype[which(landscape_list$fitness==max(landscape_list$fitness))]
mu <- 0.0001#1e-4 # mutation rate as in Szendro et al. 2013 
m <- nrow(gt_lst) # of genotypes
M<-mutation_matrix(gt_lst,mu)
S<-selection_matrix(landscape_list$fitness)
E<-M %*% S

# gstart<-paste(c(0,0,0,0,1),collapse = "_")
# gend<-paste(c(1,1,1,1,1),collapse = "_")
adaptation<-function(landscape_list,gstart,gend,verbose){
  g0i<-which(landscape_list$genotype==gstart)
  if(verbose){
    print(paste("starting genotype:", landscape_list$genotype[g0i]))
    print(paste("starting fitness:", landscape_list$fitness[g0i]))
  }
  g1i<-which(landscape_list$genotype==gend)
  
  #p is the frequencies vector
  p<-matrix(NA, nrow = nrow(landscape_list), ncol = generations)
  p[,1]<-0
  p[g0i,1]<-1.0
  
  # pom holds the index of the most frequent genotype at each time point
  pom<-matrix(NA, nrow = nrow(landscape_list), ncol = 1)
  pom[1]<-which(p[,1]==max(p[,1]))
  
  # W holds the population mean fitness at each time point
  W <- matrix(NA, nrow = generations, ncol = 1)
  W[1] = diag(S) %*% p[,1]  # mean fitness calculation
  
  for (kk in 2:generations){
    if((verbose)&(kk %% c(1000) == 0)){
      print(kk, end=" ")
    }
    p[,kk] <- E %*% p[, kk - 1]  # selection and mutation are deterministic
    p[,kk] <- p[, kk]/sum(p[,kk])  # normalize frequencies vector
    p[,kk] <- unlist(lapply(p[,kk],rbinom,n=1,size =Npop))/Npop  # random genetic drift is stochastic

    W[kk] <- diag(S) %*% p[,kk]   # mean fitness calculation
    pom[kk]<-which(p[,kk]==max(p[,kk]))
    
    # break if reached fixation
    if(max(p[g1i,kk]) > 0.99){ #<---------------------------should be adjusted according to mu
        break
    }
  }
  # return(list(kk, p, W,pom[kk]))
  return(pom[kk])
}

```

## 6.2. simulation of evolution base on the NK model
```{r Data_Source,warning=FALSE, message=FALSE,fig.height=2,fig.width=2,dpi=300}
Npop<-1e5
generations<-30 # 2^12
intrcpt<-c(0.3,0.5,0.99)
N<-10 # number of genetic sites
K<-3
Nland<-10
mu <- 0.001#1e-4 # mutation rate as in Szendro et al. 2013 
exlevel<--0.1
# Nini<-100 # activate if start from random genotypes

gt_lst<-permutations(2,N,v=c(0,1),repeats.allowed=TRUE)
m <- nrow(gt_lst) # of genotypes
NK_exlandscape_list<-NK_exlandscape(gt_lst,Nland,modify=FALSE)
NK_landscape_list<-fitlandscape(NK_exlandscape_list,intrcpt)
M<-mutation_matrix(gt_lst,mu)

gt_initials<-initial_genotypes(gt_lst,NK_exlandscape_list,exlevel)
gstart_lst<-list()
for(i in 1:length(gt_initials)){
  temp<-gt_initials[[i]]
  Nini<-min(c(length(temp),600))
  gstart_lst[[i]]<-sample(x=temp,size=Nini)
}

for (i in 1:length(NK_landscape_list)){
  landscape_list<-NK_landscape_list[[i]]
  gend<-landscape_list$genotype[which(landscape_list$fitness==max(landscape_list$fitness))]
  S<-selection_matrix(landscape_list$fitness)
  E<-M %*% S
  l<-ifelse(i%%Nland==0,Nland,i%%Nland)
  gstart<-gstart_lst[[l]]
  res<-mclapply(gstart,adaptation,landscape_list = landscape_list,gend = gend,verbose=FALSE)
  res<-unlist(res)
  temp<-createEmptyDf(nrow=length(res),ncol=4,colnames=qw(intrcpt,expression,muteffect,shape))
  temp$intrcpt<-landscape_list$intercept[1]
  temp$expression<-landscape_list$expression[res]
  temp$muteffect<-landscape_list$muteffect[res]
  temp$shape<-landscape_list$shape[res]
  if(i==1){
    dMutEffect_nk<-temp
  }else{
    dMutEffect_nk<-rbind(dMutEffect_nk,temp)
  }
  print(i)
}
```

## 6.3. Visualization
### 6.3.1. 
```{r Data_Source,warning=FALSE, message=FALSE,fig.height=3.2,fig.width=9.6,dpi=300}
dMutEffect_nk$intrcpt<-factor(dMutEffect_nk$intrcpt,levels=intrcpt)
temp<-get_palette("lancet",2)
temp<-c(temp[c(2,1)],"gray40")
myslope<--1
myintercept<-0.5
ggscatter(dMutEffect_nk,x="expression",y="muteffect",color = "intrcpt",alpha=0.2,size=0.2)

temp0<-dMutEffect_nk
tempx<-list()
figlist<-list()
for(i in 1:length(intrcpt)){
  tempx[[i]]<-temp0[temp0$intrcpt==intrcpt[i],]
  tempx[[i]]<-tempx[[i]][order(tempx[[i]]$expression),]
  tempx[[i]]$Runmed<-runmed(tempx[[i]]$muteffect,101)
  g<-ggscatter(tempx[[i]],x="expression",y="muteffect",color = temp[i],alpha=0.1,size=0.2,xlim=c(0,1),ylim=c(-0.3,0.3),
               xlab="Expression level [a.u.]", ylab="Mutational effect [a.u.]")+
    scale_x_continuous(breaks = get_breaks(by = 0.5, from = 0),limits =  c(0, 1)) +
    geom_line(aes(x=expression,y=Runmed),color = temp[i])+
    annotate("text",x=Inf,y=Inf,label=paste("Fb=",intrcpt[i],sep=" "), hjust = 1,vjust = 1,size=7)
  g<-ggpar(g,font.xtickslab = c(18),font.ytickslab = c(18),font.x = c(18),font.y=c(18))
  # g<-g+geom_abline(slope = myslope,intercept = myintercept,size=0.2)+geom_vline(xintercept = c(0.5,0.75),size=0.2)
  figlist[[i]]<-g
}
grid.arrange(figlist[[1]], figlist[[2]], figlist[[3]], nrow = 1)

for(i in 1:length(intrcpt)){
  if(i==1){
    dfrunmed_nk<-tempx[[i]]
  }else{
    dfrunmed_nk<-rbind(dfrunmed_nk,tempx[[i]])
  }
}

```
### 6.3.2. Running median
```{r Data_Source,warning=FALSE, message=FALSE,fig.height=3.2,fig.width=3.2,dpi=300}
temp1<-NK_exlandscape_list[[1]]
temp1$uphillratio<-uphill_ratio(temp1$shape)
g0<-ggscatter(temp1,x="expression",y="muteffect",color="uphillratio",alpha=0.6,size=1,
          xlab="Expression level [a.u.]",ylab="Mutation effects [a.u.]",legend.title="",xlim=c(0,1),ylim=c(-0.3,0.3))+
          scale_color_gradient2(low  = "#C26B51",mid  = "#DDCE47",high = "#529985",midpoint = 0.5)
# g0<-g0+geom_abline(slope = myslope,intercept = myintercept,size=0.2)+geom_vline(xintercept = c(0.5,0.75),size=0.2)
g1<-ggline(dfrunmed_nk,x="expression",y="Runmed",color = "intrcpt",palette = temp,
           numeric.x.axis = TRUE,plot_type="l",xlim=c(0,1),ylim=c(-0.3,0.3),
           xlab="Expression level [a.u.]", ylab="Mutational effects [a.u.]",legend.title="")
# g1<-g1+geom_abline(slope = myslope,intercept = myintercept,size=0.2)
g1.1<-ggline(dfrunmed_nk,x="expression",y="Runmed",color = "intrcpt",palette = temp,
           numeric.x.axis = TRUE,plot_type="l",xlim=c(0.6,0.8),ylim=c(-0.2,0),
           xlab="Expression level [a.u.]", ylab="Mutational effects [a.u.]")

grid.arrange(g0,g1, nrow = 1)

g5<-ggline(dfrunmed_nk,x="expression",y="Runmed",color = "intrcpt",palette = temp,
           numeric.x.axis = TRUE,plot_type="l",xlim=c(0,1),ylim=c(-0.4,0.4),
           xlab="Expression level [a.u.]", ylab="Mutational effect [a.u.]",legend="")+
  scale_x_continuous(breaks = get_breaks(by = 0.5, from = 0),limits =  c(0, 1)) +
  scale_y_continuous(breaks = get_breaks(by = 0.2, from = -0.4),limits =  c(-0.4, 0.4))
g5<-ggpar(g5,font.xtickslab = c(18),font.ytickslab = c(18),font.x = c(18),font.y=c(18))
g5

g5e<-ggline(dfrunmed_nk,x="expression",y="Runmed",color = "intrcpt",palette = temp,
           numeric.x.axis = TRUE,plot_type="l",xlim=c(0.4,0.8),ylim=c(-0.15,0.1),
           xlab="Expression level [a.u.]", ylab="Mutational effect [a.u.]",legend="")+
  scale_x_continuous(breaks = get_breaks(by = 0.2, from = 0.4),limits =  c(0.4, 0.8)) +
  scale_y_continuous(breaks = get_breaks(by = 0.1, from = -0.2),limits =  c(-0.15, 0.1))
g5e<-ggpar(g5e,font.xtickslab = c(18),font.ytickslab = c(18),font.x = c(18),font.y=c(18))
g5e
```
### 6.3.3. 
```{r Data_Source,warning=FALSE, message=FALSE,fig.height=3.2,fig.width=9.6,dpi=300}
temp1<-NK_exlandscape_list[[1]]
temp1$uphillratio<-uphill_ratio(temp1$shape)

g1<-ggscatter(temp1,x="hmdist",y="expression",color="uphillratio",alpha=0.5,size=2,
          xlab="Hamming distance",ylab="Expression level [a.u]",legend="")+
    scale_x_continuous(breaks = get_breaks(by = 2, from = 0),limits =  c(0, 10)) +
  scale_y_continuous(breaks = get_breaks(by = 0.5, from = 0),limits =  c(0, 1)) +
  scale_color_gradient2(low  = "#C26B51",mid  = "#DDCE47",high = "#529985",midpoint = 0.5,limits=c(0,1))
g1<-ggpar(g1,font.xtickslab = c(18),font.ytickslab = c(18),font.x = c(18),font.y=c(18))

temp3<-NK_landscape_list[[1]]
temp3$uphillratio<-uphill_ratio(temp3$shape)
g2<-ggscatter(temp3,x="hmdist",y="fitness",color="uphillratio",alpha=0.5,size=2,ylim=c(0,1),
          xlab="Hamming distance",ylab="Fitness [a.u]",legend="")+
  geom_hline(yintercept = intrcpt[1],color=temp[1])+
  scale_x_continuous(breaks = get_breaks(by = 2, from = 0),limits =  c(0, 10)) +
  scale_y_continuous(breaks = get_breaks(by = 0.2, from = 0),limits =  c(0, 1)) +
  scale_color_gradient2(low  = "#C26B51",mid  = "#DDCE47",high = "#529985",midpoint = 0.5,limits=c(0,1))
g2<-ggpar(g2,font.xtickslab = c(18),font.ytickslab = c(18),font.x = c(18),font.y=c(18))

g3<-ggscatter(temp1,x="expression",y="muteffect",color="uphillratio",alpha=0.5,size=2,
          xlab="Expression level [a.u.]",ylab="Mutational effect [a.u.]",legend="",xlim=c(0,1),ylim=c(-0.4,0.4))+
          scale_x_continuous(breaks = get_breaks(by = 0.5, from = 0),limits =  c(0, 1)) +
          scale_color_gradient2(low  = "#C26B51",mid  = "#DDCE47",high = "#529985",midpoint = 0.5,limits=c(0,1))
g3<-ggpar(g3,font.xtickslab = c(18),font.ytickslab = c(18),font.x = c(18),font.y=c(18))

grid.arrange(g1,g3,g2, nrow = 1)
```
## 6.4. Output
```{r RMF,warning=FALSE, message=FALSE,fig.height=3.2,fig.width=3.2,dpi=300}
# expression landscape
for(i in 1:length(NK_exlandscape_list)){
  my_tmp<-NK_exlandscape_list[[i]]
  my_tmp$ID<-i
  if(i==1){
    my_df<-my_tmp
  }else{
    my_df<-bind_rows(my_df,my_tmp)
  }
}
my_df<-my_df %>% select(ID,everything())
write_csv(my_df,file = paste(getwd(),"Output","ExprsLandscape_NK.csv",sep = "/"))

# fitness landscape
my_ids<-rep(seq(1,length(NK_exlandscape_list)),each=length(intrcpt))
for(i in 1:length(NK_landscape_list)){
  my_tmp<-NK_landscape_list[[i]]
  my_tmp$ID<-my_ids[i]
  if(i==1){
    my_df<-my_tmp
  }else{
    my_df<-bind_rows(my_df,my_tmp)
  }
}
my_df<-my_df %>% select(ID,everything())
write_csv(my_df,file = paste(getwd(),"Output","FitnessLandscape_NK.csv",sep = "/"))

# mutational effect
write_csv(dMutEffect_nk,file = paste(getwd(),"Output","MutEffect_NK.csv",sep = "/"))
```